A Survey on Ordered Binary Decision Diagrams

### 2015-09-15 paper.tex
Revise Implementational aspects
Revise Efficient data structures

Hence, the following sections cover important aspects to keep in mind when
implementing an OBDD framework efficiently, i.e. efficient data structures, the
$ITE$ operator, how to possibly approach the variable ordering problem and
\textit{garbage collection}. \tbr

@@

Hash tables maintain \textit{key}-\textit{value}-pairs, returning the associated
value for a given key in (amortized) constant time. Collisions, i.e. equivalent
value for different keys, are resolved by chaining all associated values in a 
linked list~\cite{BRACE90}.

This is especially useful to refine \textit{Apply} in order to guarantee the time
complexity of $O\left(|V_{f_1}| \cdot |V_{f_2}|\right)$. Using a hash table to
match pairs of nodes ($n_1$, $n_2$) to a single node $m$, \textit{Apply} avoids
making multiple recursive calls to equivalent pairs of nodes and thus avoids
computing equivalent subgraphs multiple times. Therefore, before computing a
subgraph, \textit{Apply} checks whether an entry is present for the given pair
of nodes. If so, the associated value $m$ is used instead of computing the
subgraph~\cite{BRYANT86, BRYANT92}.

Moreover, hash tables are useful to generate a reduced graph directly from some
operations by maintaing a hash table associating triples
($var(n)$, $lo(n)$, $hi(n)$) with a single nonterminal node $n$. This is done for
each nonterminal node $n$ generated while executing the
operation~\cite{BRYANT92}. This is also referred to as
\textit{unique-table}~\cite{BRACE90} imposes the canonical form of an OBDD. If a
new node is generated, a corresponding entry is created in the unique-table.

There exists a special version of hash tables, so-called
\textit{hash-based caches}, which do not resolve collisions, i.e. overwrite
existing entries, and issue a cache miss if a looked up element is not found.
As will be seen, \textit{Hash-based caches} are useful to provide an efficient
implementation of the $ITE$ operator. \tbr

@@

To reduce the number of nodes of an OBDD, one can use \textit{complement edges}.
A complement edge is an edge holding a single bit to indicate complementation.
Hence, when $\overline{f}$ needs to be represented and a node $f$ is already
present, a complement edge can be used to connect $f$ and thus avoid
constructing the intermediate node $\overline{f}$. Also one of the two terminal
nodes (e.g. \textbf{1}) can be removed by just using a complement edge to connect
the other one (e.g. \textbf{0})~\cite{BRACE90}.

Brace et al.~\cite{BRACE90} identified four functionally equivalent pairs of
functions to be used for complement edges in order to maintain the canonical form.
Let $x$ be the node having $lo(x)$ and $hi(x)$ on its outgoing 0- and 1-edge,
respectively. Moreover, let $in(x)$ be the (one) incoming edge of node $x$ and
$\overline{c}$ the complement of $c$. Then, the following table sums up the
four pairs identified by Brace et al.~\cite{BRACE90}.

@@

Also, the 1-edge of each node must always be a regular one to guarantee a
canonical form.

Implementing complement edges enables one to compute the complement and identify
complement functions in constant time (because $f$ and $\overline{f}$ use the same
node). According to Brace et al.~\cite{BRACE90} this reduces the total runtime
to form the OBBDD by almost a factor of 2 (for 12 examples).

Another refinement to the graph representation, also proposed by Brace et
al.~\cite{BRACE90}, is to integrate the unique-table into the graph representation,
i.e. linking the collision chain of each node to the corresponding unique-table
entry, to improve memory usage. \tbr

>>>>

Hence, the following sections cover important aspects to keep in mind when
implementing an OBDD framework efficiently, i.e. efficient data structures, the
$ITE$ operator, \textit{garbage collection} and one possible approach to the variable
ordering problem.

@@

Hash tables maintain \textit{key}-\textit{value}-pairs, returning the associated
value for a given key in (amortized) constant time. Collisions, i.e. equivalent
values for different keys, are resolved by chaining all associated values in a 
linked list~\cite{BRACE90}.

This is especially useful to refine \textit{Apply} in order to guarantee the time
complexity of $O\left(|V_{f_1}| \cdot |V_{f_2}|\right)$. Using a hash table to
match pairs of nodes ($n_1$, $n_2$) to a single node $m$, \textit{Apply} avoids
making multiple recursive calls to equivalent pairs of nodes and thus avoids
computing equivalent subgraphs multiple times. Therefore, before computing a
subgraph, \textit{Apply} checks whether an entry is present for the given pair
of nodes. If so, the associated value $m$ is used instead of computing the
subgraph again~\cite{BRYANT86, BRYANT92}.

Moreover, hash tables are useful to generate a reduced graph directly from some
operations by maintaing a hash table associating triples
($var(n)$, $lo(n)$, $hi(n)$) with a single nonterminal node $n$. This is done for
each nonterminal node $n$ generated while executing the
operation~\cite{BRYANT92}. This is also referred to as
\textit{unique-table}~\cite{BRACE90} and imposes the canonical form of an OBDD. If a
new node is generated, a corresponding entry is created in the unique-table.

There exists a special type of hash tables, so-called
\textit{hash-based caches}, which do not resolve collisions, i.e. they overwrite
existing entries and issue a \textit{cache miss} if a looked up element is not
found. As will be seen, \textit{hash-based caches} are useful to provide an
efficient implementation of the $ITE$ operator. \tbr

@@

To reduce the number of nodes of an OBDD, one can use \textit{complement edges}.
A complement edge is an edge holding a single bit to indicate complementation.
Hence, when $\overline{f}$ needs to be represented and a node $f$ is already
present, a complement edge can be used to connect $f$ and thus avoid
constructing the intermediate node $\overline{f}$. Using this, one of the two
terminal nodes (e.g. \textbf{1}) can be removed too, by using a complement edge
to connect the other one (e.g. \textbf{0})~\cite{BRACE90}.

Brace et al.~\cite{BRACE90} identified four functionally equivalent pairs of
functions to be used for complement edges in order to maintain the canonical form.
Let $x$ be the node having $lo(x)$ and $hi(x)$ on its outgoing 0- and 1-edge,
respectively. Moreover, let $in(x)$ be the (one) incoming edge of node $x$ and
$\overline{c}$ be the complement of $c$. In the following, these four pairs,
identified by Brace et al.~\cite{BRACE90}, are summarized:

@@

Also, the 1-edge of each node must always be a regular one to guarantee the
canonical form.

Implementing complement edges enables one to compute the complement and identify
complement functions in constant time (because $f$ and $\overline{f}$ use the same
node). According to Brace et al.~\cite{BRACE90}, this technique reduces the total
runtime to form the OBDD by almost a factor of 2 (for 12 examples).

Another refinement to the graph representation, also proposed by Brace et
al.~\cite{BRACE90}, is to integrate the unique-table into the graph representation,
i.e. linking the collision chain of each node to the corresponding unique-table
entry, in order to improve memory usage.
